---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# epts

<!-- badges: start -->
[![R-CMD-check](https://github.com/george-wood/epts/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/george-wood/epts/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![test-coverage](https://github.com/george-wood/epts/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/george-wood/epts/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

epts is a package for reading [Electronic Performance and Tracking Systems](https://www.fifa.com/technical/football-technology/standards/epts) (EPTS) data.

## Overview

EPTS is a broad term for camera-based and wearable technologies that measure and track football (soccer) players. This includes positioning systems that track player and ball positions, accelerometers, gyroscopes, and other technologies. Such data is widely used in sports analytics.

FIFA and FC Barcelona have developed a standard format for exchanging EPTS data. The standard format requires data providers to produce two documents when exchanging data:

- A raw data file, which contains the actual data in a format that can be parsed line by line
- An xml file containing metadata and a data format specification that defines the format of the raw data file

The epts package makes it easy to read the raw data file according to the data format specification.

## Features

With epts, you can:

- Import the raw data file according to the data format specification
- Extract metadata

When importing raw data, epts returns a `data.frame` with column names as defined in the data format specification. For flexibility, the metadata is extracted as a `list`.

## Installation

You can install the development version of epts from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("george-wood/epts")
```

## Usage

```{r library}
library(epts)
```

The package includes [example data](https://www.fifa.com/technical/football-technology/standards/epts/research-development-epts-standard-data-format) provided by FIFA:

```{r example}
# player and ball tracking data
raw_data <- epts_example("fifa_example.txt")
readLines(raw_data, warn = FALSE)

# xml file including metadata and format specification
metadata <- epts_example("fifa_example.xml")
```

Import the raw data using the format specification:

```{r read}
read_epts(raw_data, metadata)
```
