---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# epts

<!-- badges: start -->
[![R-CMD-check](https://github.com/george-wood/epts/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/george-wood/epts/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

epts is a package for reading data that follows the [Electronic Performance and Tracking Systems (EPTS) standard data format](https://www.fifa.com/technical/football-technology/standards/epts/research-development-epts-standard-data-format) into R.
It provides functions to parse metadata and data format specifications, as provided in an `.xml` file. It uses these specifications to imports the raw data `.txt` file and return a data frame. 

The package makes it easy to import any EPTS data that follows the standard format into R.

## Installation

You can install the development version of epts from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("george-wood/epts")
```

## Usage

This example imports the [example data provided by FIFA](https://www.fifa.com/technical/football-technology/standards/epts/research-development-epts-standard-data-format):

```{r example}
library(epts)

# player and ball tracking data
raw_data <- epts_example("fifa_example.txt")

# metadata and data format specification file
metadata <- epts_example("fifa_example.xml")

# import the example data
read_epts(raw_data, metadata)
```
